---
title: "The Promise and Perils of Packaging Your Processes"
subtitle: "(but mostly promise)"
author: 'Karl Hailperin <br> <a href="http://github.com/khailper"><i class="fa fa-github"></i>&nbsp; khailper</a><br> <a href="http://twitter.com/khailper"> <i class="fa fa-twitter"></i> &nbsp; @khailper</a><br>'
date: "2018/06/20"
output:
  xaringan::moon_reader:
    css: ["default", "default-fonts", "footer.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

layout: true
```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

<div class="my-footer"><span>khailper.github.io/process_packaging_pres</span></div>
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">

<!-- reusing footer code from https://github.com/rstudio-education/arm-workshop-rsc2019/blob/master/static/slides/xaringan.Rmd, https://github.com/rstudio-education/arm-workshop-rsc2019/blob/master/static/slides/css/sfah.css, and https://github.com/yihui/xaringan/wiki/Footer-and-header-lines stylesheet code from https://holtzy.github.io/Pimp-my-rmd/#footer_and_header-->

---
# Who am I?

- Data science consultant
- Have built R packages for a client to streamline data processing and 
visualization
- They/them pronouns

---
# What this talk is (and isn't) about
- Not (necessarily) about destined for Github/CRAN
- Not about how to build packages/intro to `devtools` and `roxygen2` (but there 
will be resources on that at the end)
- It's about practices I've found useful in developing packages for personal use 
or at work.

---
# Why should I bother building a package if no one else is going to use it?
- `devtools` and `usethis` provide tools for documenting what you were 
thinking when you wrote the code and what your dependencies are
- `testthat` makes it easier to build unit tests that will keep you from 
breaking things (or at least not noticing that you're breaking things)
- Future you will be grateful
- Low-stress way to develop related skills
- Good rule of thumb (that I can't find the source for): If you re-use the same 
code three times, write a function. If you re-use the same function three times,
write a package.

---
# Also, you might be sharing your processes with your co-workers
- shared != public
- It's likely your team is going to run into same problem. An internal package 
means you only have to solve it once.
- Code review is quicker when you don't have to double-check the same things
every time.
- Using a package of common processes also acts as a style guide.

---
# How repeated code is like for loops 
## (and therefore cupcakes)
- [For those unfamiliar with the above reference](https://www.youtube.com/watch?v=GyNqlOjhPCQ)
- Repeating code obscures changes.
- Repeating code leads to accidental changes.
- Need to remember to manually propagate changes to all versions.

---
# utils.R is your friend (internal functions)
- Remember that not everything needs to be `@exported`-ed
- moving chunks of code into a (well-named) internal function can make it easier
to review and test your code
- "Rule of three" applies to code within functions
- makes it easy to update a common workflow or use a parameter to 'switch' 
which workflow your using

---
# Using sensible defaults
- Use defaults to hide repeated processes that you don't always want to run.
- For me, this was driven home by a _very_ slow processing step I almost never 
used.
- often a simple TRUE/FALSE parameter is enough

---
# Case study in internal functions and defaults
- Starting point was something like this:
```{r basic_pseudo_code}
# assorted documentation ommitted
# usethis::use_pipe() allows us to use %>% 
#' @export
read_page_views <- function(){
  here::here("data", "page_views.csv") %>% 
  readr::read_csv() %>% 
  # data comes in as UTC
  dplyr::mutate(time = lubridate::with_tz(time, 
                                          tz = "America/Chicago"),
                # some non-standard characters in page names were 
                # coming out garbled, and needed to be replaced 
                # for use in plot labels
                page_name = regex_stuff(page_name))
}
```

---
# But then we added bunch of other data sources...
--

- Also, there's discussion of reading directly from AWS.
--

- Now code looks more like `here::here("data", "web_data","page_views.csv")`.
--

- So now `read_page_views()` looks more like this:
--

```{r use_read_data_file}
#' @export
read_page_views <- function(){
  read_data_file(location = "web_data", #<<
                 file_name = "page_views.csv") %>% #<<
  # data comes in as UTC
  dplyr::mutate(time = lubridate::with_tz(time, 
                                          tz = "America/Chicago"),
                page_name = regex_stuff(page_name))
}
```
--

- `read_data_file` is an internal function that wraps the first two steps

---
# That seems overly complicated...
--

- That's not entirely unfair.
--

- I may have omitted some code.
--

- The code looked more like this:
```{r location}
#' @export
read_page_views <- function(location){
  data_file <- dplyr::case_when( #<<
    location == "local" ~ read_data_file(location = "web_data", #<<
                                         file_name = "page_views.csv"), #<<
    location == "aws" ~ read_data_file(location = "aws", #<<
                                         file_name = "page_views.csv"), #<<
    TRUE ~ stop("Helpful error message") #<<
  ) #<<
  data_file %>% 
  # data comes in as UTC
  dplyr::mutate(time = lubridate::with_tz(time, 
                                          tz = "America/Chicago"),
                page_name = regex_stuff(page_name))
}
```

---
# read_data_file() pseudo-code
```{r read_data_file}
read_data_file <- function(location, file_name){
  if (location = "aws"){
    # code to handle reading from AWS
  } else{
    here::here("data", location, file_name) %>% 
      readr::read_csv()
  }
}
```
- Added bonus: we can give a default value for `location` parameters.
- Went from `read_page_views <- function(location = "local"){# ...}` to
`read_page_views <- function(location = "aws"){# ...}` when AWS 
authorization system was ready.
- Actual code still just uses `read_page_views()`.
- Plus, `read_data_file()` makes it easy to create new `read_*()` functions as
we add data sources. Just edit the wrapper and write some data-specific unit 
tests.

---
# Remember regex_stuff(page_name)?

--

- Yes...why?
--

- It turns out it takes *forever* and almost never matters
--

- Better implementation:
--

```{r procces_name}
#' @export
read_page_views <- function(location = "local", process_name = FALSE){
  data_file <- dplyr::case_when(
    location == "local" ~ read_data_file(location = "web_data",
                                         file_name = "page_views.csv"),
    location == "aws" ~ read_data_file(location = "aws",
                                         file_name = "page_views.csv"),
    TRUE ~ stop("Helpful error message")
  ) %>% 
    mutate(time = lubridate::with_tz(time, 
                                     tz = "America/Chicago"))
  if (process_name){ #<<
    dplyr::mutate(data_file, #<<
                  page_name = regex_stuff(page_name)) #<<
  } else { #<<
    return(data_file) #<<
  } #<<
}
```

---
# Additional resources
- [RStudio's cheetsheet](https://github.com/rstudio/cheatsheets/blob/master/package-development.pdf)
- [_R Packages_ by Hadley Wickham](http://r-pkgs.had.co.nz/)
- [WIP 2<sup>nd</sup> edition with Jenny Bryan](https://r-pkgs.org/)
- ["Writing an R package from scratch" by Hilary 
Parker](https://hilaryparker.com/2014/04/29/writing-an-r-package-from-scratch/)
- [Thomas Westlake's version of "Writing an R package from scratch" using 
`usethis`](https://r-mageddon.netlify.com/post/writing-an-r-package-from-scratch/)
- ["R Package Primer" by Karl Broman](https://kbroman.org/pkg_primer/)
- ["`usethis` workflow for package development" by Emil 
Hvitfedlt](https://www.hvitfeldt.me/blog/usethis-workflow-for-package-development/)
- [The tidyverse style guide](https://style.tidyverse.org/)

---
# Thank you/questions
